Simulation Tables

### 2019-07-09 manuscript/manuscript.tex
sim 1

For cluster $k$, we collected all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{X}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ to utilize the matrix normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$. For the multinomial regression model component of this simulation, we modeled the class labels $z_i$ as a function of an intercept and one baseline covariate, $\text{w}_{i1}$, implying that $r = 2$. Since a comparison of imputation method to alternative approaches is performed later in our second simulation study, we did not introduce missing data into this simulatation. As a result, the final sample size was $N = n \times J = 4,000$.

We chose the MSN hyperparameters to be homogeneous across the three clusters by setting, for $k = 1,2,3$, $\mathbf{B}_{0k}^* = \mathbf{0}_{4 \times 3}$, $\mathbf{V}_{0k} = \mathds{1}_{4 \times 4}$, and $\nu_{0k} = J + 2 = 6$, which gives $\text{E}(\boldsymbol\Sigma_k) = \mathds{1}_{4 \times 4}$. Similarly, we set $\mathbf{d}_{01} = \mathbf{d}_{02} = (0,0)^T$ and $\mathbf{S}_{01} = \mathbf{S}_{02} = \mathds{1}_{2 \times 2}$, noting that $k = 3$ is the reference cluster. To investigate the effect of ignoring skewness, we compare MSN and MVN models in terms of parameter estimates and WAIC as described below.

@@

\caption{\label{tab:unnamed-chunk-5}Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2. 1,000 iterations were run with a burn in of 100. Missingness mechanism was MAR and P(miss) = 0. Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}{lllll}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Class 1} \\
\cmidrule(l{3pt}r{3pt}){3-5}
Component & Param. & True & MSN Est. (95\% CrI) & MN Est. (95\% CrI) \\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{ }}\\
\hspace{1em}MVSN & $\beta_{11}$ & 11 & 11.07 (10.74, 11.39) & 9.42 (8.91, 9.77)\\
\hspace{1em}Regression & $\beta_{21}$ & 12 & 12.02 (11.87, 12.17) & 11.98 (11.77, 12.18)\\
\hspace{1em} & $\beta_{31}$ & 13 & 13.06 (12.75, 13.36) & 11.39 (10.7, 11.78)\\
\hspace{1em} & $\beta_{41}$ & 14 & 14.06 (13.91, 14.22) & 14.02 (13.78, 14.22)\\
\hspace{1em} & $\beta_{12}$ & 2 & 2.11 (1.82, 2.35) & 0.42 (0.03, 0.83)\\
\hspace{1em} & $\beta_{22}$ & 2 & 2.03 (1.88, 2.17) & 2.02 (1.86, 2.22)\\
\hspace{1em} & $\beta_{32}$ & 2 & 2.13 (1.8, 2.43) & 0.49 (0.14, 0.86)\\
\hspace{1em} & $\beta_{42}$ & 2 & 2.08 (1.93, 2.23) & 2.08 (1.92, 2.28)\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{ }}\\
\hspace{1em} & $\alpha_{1}$ & -0.99 & -0.81 (-2.12, 0.05) &     / \\
\hspace{1em} & $\alpha_{2}$ & -0.5 & -0.22 (-1.3, 0.75) &     / \\
\hspace{1em} & $\alpha_{3}$ & -0.5 & -0.96 (-2.14, 0.01) &     / \\
\hspace{1em} & $\alpha_{4}$ & -0.99 & -1.18 (-2.44, -0.06) &     / \\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{ }}\\
\hspace{1em}Multinom. & $\delta_{11}$ & -0.08 & -0.07 (-0.27, 0.12) & -0.54 (-0.77, -0.32)\\
\hspace{1em} & $\delta_{12}$ & 0.51 & 0.25 (-0.04, 0.53) & -0.26 (-0.6, 0.05)\\
\hspace{1em} & $\delta_{21}$ & -0.97 & -0.91 (-1.15, -0.68) & -0.07 (-0.28, 0.14)\\
\hspace{1em} & $\delta_{22}$ & 0.84 & 0.39 (0.09, 0.71) & 0.24 (-0.04, 0.5)\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{ }}\\
\hspace{1em}Clustering & $\pi_l$ & 0.38 & 0.38 (0.38, 0.38) & 0.38 (0.13, 0.41)\\
\bottomrule
\end{tabular}
\end{table}

>>>>

For cluster $k$, we collected all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{X}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ to utilize the matrix normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$. For the multinomial regression model component of this simulation, we modeled the class labels $z_i$ as a function of an intercept and one baseline covariate, $\text{w}_{i1}$, implying that $r = 2$. We did introduce missing data into this simulation, as missing data will be dealt with in our second simulation study. As a result, the final sample size was $N = n \times J = 4,000$.

We chose the MSN hyperparameters to be homogeneous across the three clusters by setting, for $k = 1,2,3$, $\mathbf{B}_{0k}^* = \mathbf{0}_{4 \times 3}$, $\mathbf{V}_{0k} = \mathds{1}_{4 \times 4}$, and $\nu_{0k} = J + 2 = 6$, which gives $\text{E}(\boldsymbol\Sigma_k) = \mathds{1}_{4 \times 4}$. Similarly, we set $\mathbf{d}_{01} = \mathbf{d}_{02} = (0,0)^T$ and $\mathbf{S}_{01} = \mathbf{S}_{02} = \mathds{1}_{2 \times 2}$, noting that $k = 3$ is the reference cluster. To investigate the effect of ignoring skewness, we fit both the MVN and MSN mixture models to data generated from model (\ref{eq:sim1}). The values of WAIC for the MSN mixture model and MVN mixture model were 37,221.7 and 324,705.3, respectively, indicating model fit suffered when skewness was ignored. Furthermore, we report below in Table \ref{tab:sim1_parms} a selection of parameter estimates from the MSN and MVN models that show incorrect inference (i.e., true parameter values not contained in 95\% posterior credible intervals) resulted from ignoring skewness. A full table showing all model parameters estimates and 95\% posterior credible intervals for the MSN and MVN models in addition to ground truth values is given in Web Appenxix B.

@@

\caption{\label{tab:sim1_parms}Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2. 1,000 iterations were run with a burn in of 100. Missingness mechanism was MAR and P(miss) = 0. Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}{lllll}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Class 1} \\
\cmidrule(l{3pt}r{3pt}){3-5}
Component & Param. & True & MSN Est. (95\% CrI) & MN Est. (95\% CrI) \\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{ }}\\
\hspace{1em}MVSN & $\beta_{11}$ & 11 & 11.07 (10.74, 11.39) & 9.42 (8.91, 9.77)\\
\hspace{1em}Regression & $\beta_{21}$ & 12 & 12.02 (11.87, 12.17) & 11.98 (11.77, 12.18)\\
\hspace{1em} & $\beta_{31}$ & 13 & 13.06 (12.75, 13.36) & 11.39 (10.7, 11.78)\\
\hspace{1em} & $\beta_{41}$ & 14 & 14.06 (13.91, 14.22) & 14.02 (13.78, 14.22)\\
\hspace{1em} & $\beta_{12}$ & 2 & 2.11 (1.82, 2.35) & 0.42 (0.03, 0.83)\\
\hspace{1em} & $\beta_{22}$ & 2 & 2.03 (1.88, 2.17) & 2.02 (1.86, 2.22)\\
\hspace{1em} & $\beta_{32}$ & 2 & 2.13 (1.8, 2.43) & 0.49 (0.14, 0.86)\\
\hspace{1em} & $\beta_{42}$ & 2 & 2.08 (1.93, 2.23) & 2.08 (1.92, 2.28)\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{ }}\\
\hspace{1em} & $\alpha_{1}$ & -0.99 & -0.81 (-2.12, 0.05) &     / \\
\hspace{1em} & $\alpha_{2}$ & -0.5 & -0.22 (-1.3, 0.75) &     / \\
\hspace{1em} & $\alpha_{3}$ & -0.5 & -0.96 (-2.14, 0.01) &     / \\
\hspace{1em} & $\alpha_{4}$ & -0.99 & -1.18 (-2.44, -0.06) &     / \\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{ }}\\
\hspace{1em}Multinom. & $\delta_{11}$ & -0.08 & -0.07 (-0.27, 0.12) & -0.54 (-0.77, -0.32)\\
\hspace{1em} & $\delta_{12}$ & 0.51 & 0.25 (-0.04, 0.53) & -0.26 (-0.6, 0.05)\\
\hspace{1em} & $\delta_{21}$ & -0.97 & -0.91 (-1.15, -0.68) & -0.07 (-0.28, 0.14)\\
\hspace{1em} & $\delta_{22}$ & 0.84 & 0.39 (0.09, 0.71) & 0.24 (-0.04, 0.5)\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{ }}\\
\hspace{1em}Clustering & $\pi_l$ & 0.38 & 0.38 (0.38, 0.38) & 0.38 (0.13, 0.41)\\
\bottomrule
\end{tabular}
\end{table}

An attractive feature of the MSN model is that it includes the MVN model as a special case. As such, the MSN model can be used in place of the MVN model even when data are not clearly skewed. We demonstrate this by fitting both the MSN and MVN model to data generated from model (\ref{eq:sim1}), where all skewness parameters are set to zero. As shown Table 2 of Web Appendix B, parameter estimates are similar between the MSN and MVN models, and all 95\% credible intervals for the $\alpha$ parameters of the MSN models contain 0, indicating the MSN model performs well even when data are not skewed.

### 2019-07-09 mcmc_draws_2019-07-09_SIM1/mcmc_draws_NOSKEW_2019-07-09/tables.Rmd
sim 1

```{r}
setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02/mcmc_draws_NOSKEW_2019-07-02/")
load("BETA")
load("beta_true")
load("DELTA")
load("delta_true")
load("PI")
load("PSI")
load("psi_true")
load("SIGMA")
load("sig2_true")
load("pi_true")
load("Z")
```

>>>>

```{r}
setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09/mcmc_draws_NOSKEW_2019-07-09/")
load("BETA")
load("beta_true")
load("DELTA")
load("delta_true")
load("PI")
load("PSI")
load("psi_true")
load("SIGMA")
load("sig2_true")
load("pi_true")
load("Z")
```

### 2019-07-09 mcmc_draws_2019-07-09_SIM1/tables.Rmd
sim 1

```{r}
setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-08")
load("BETA")
load("beta_true")
load("DELTA")
load("delta_true")
load("PI")
load("PSI")
load("psi_true")
load("SIGMA")
load("sig2_true")
load("pi_true")
load("Z")
```

@@

load("mcmc_draws_NOSKEW_2019-07-02/class_1_est")
load("mcmc_draws_NOSKEW_2019-07-02/class_2_est")
load("mcmc_draws_NOSKEW_2019-07-02/class_3_est")

t1 <- cbind(model_component,
            params,
            true_c1,est_c1,est_c3_noskew,
            true_c2,est_c2,est_c2_noskew,
            true_c3,est_c3,est_c1_noskew)
colnames(t1) <- c("Component",
                  "Param.",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI) ",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)")
```

```{r}
rownames(t1) <- NULL
kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1000, k = 4, p = 2, h = 3, r = 2. 1000 iterations were run with a burn in of 100. Missingness mechanism was MAR and P(miss) = 0. Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.") %>%
  kable_styling(font_size = 7) %>%
  add_header_above(c(" " = 2,"Class 1" = 3, "Class 2" = 3, "Class 3" = 3)) %>%
  group_rows(" ",1,8) %>%
  group_rows(" ",9,18) %>%
  group_rows(" ",19,22) %>%
  group_rows(" ",23,26) %>%
  group_rows(" ",27,27) %>%
  landscape()
```

>>>>

```{r}
setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09")
load("BETA")
load("beta_true")
load("DELTA")
load("delta_true")
load("PI")
load("PSI")
load("psi_true")
load("SIGMA")
load("sig2_true")
load("pi_true")
load("Z")
```

@@

load("mcmc_draws_NOSKEW_2019-07-09/class_1_est")
load("mcmc_draws_NOSKEW_2019-07-09/class_2_est")
load("mcmc_draws_NOSKEW_2019-07-09/class_3_est")

t1 <- cbind(model_component,
            params,
            true_c1,est_c1,est_c2_noskew,
            true_c2,est_c2,est_c1_noskew,
            true_c3,est_c3,est_c3_noskew)
colnames(t1) <- c("Component",
                  "Param.",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI) ",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)")
```

```{r}
rownames(t1) <- NULL
kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2. 1,000 iterations were run with a burn in of 100. Missingness mechanism was MAR and P(miss) = 0. Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.") %>%
  kable_styling(font_size = 7) %>%
  add_header_above(c(" " = 2,"Class 1" = 3, "Class 2" = 3, "Class 3" = 3)) %>%
  group_rows(" ",1,8) %>%
  group_rows(" ",9,18) %>%
  group_rows(" ",19,22) %>%
  group_rows(" ",23,26) %>%
  group_rows(" ",27,27) %>%
  landscape()
```

### 2019-07-09 mcmc_draws_2019-07-09_SIM1/tables_short_sim1.Rmd
sim 1

model_component <- c("MVSN","Regression","","","","","","","","","","","","","","","","","","","","",
                     "Multinom.","","","",
                     "Clustering")
params <- c("$\\beta_{11}$","$\\beta_{21}$","$\\beta_{31}$","$\\beta_{41}$",
            "$\\beta_{12}$","$\\beta_{22}$","$\\beta_{32}$","$\\beta_{42}$",
            "$\\Omega_{11}$","$\\Omega_{12}$","$\\Omega_{13}$","$\\Omega_{14}$",
            "$\\Omega_{22}$","$\\Omega_{23}$","$\\Omega_{24}$",
            "$\\Omega_{33}$","$\\Omega_{34}$",
            "$\\Omega_{44}$",
            "$\\alpha_{1}$","$\\alpha_{2}$", "$\\alpha_{3}$","$\\alpha_{4}$",
            "$\\delta_{11}$","$\\delta_{12}$",
            "$\\delta_{21}$","$\\delta_{22}$",
            "$\\pi_l$")

true_c1 <- c(beta_true_c1,omega_true_c1,alpha_true_c1,delta_true,pi_true[1])
true_c2 <- c(beta_true_c2,omega_true_c2,alpha_true_c2,delta_true,pi_true[2])
true_c3 <- c(beta_true_c3,omega_true_c3,alpha_true_c3,delta_true,pi_true[3])
est_c1 <- c(beta_est_c1,omega_est_c1,alpha_est_c1,delta_est,pi_est[1])
est_c2 <- c(beta_est_c2,omega_est_c2,alpha_est_c2,delta_est,pi_est[2])
est_c3 <- c(beta_est_c3,omega_est_c3,alpha_est_c3,delta_est,pi_est[3])

@@

t1 <- cbind(model_component,
            params,
            true_c1,est_c1,est_c2_noskew,
            true_c2,est_c2,est_c1_noskew,
            true_c3,est_c3,est_c3_noskew)
colnames(t1) <- c("Component",
                  "Param.",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI) ",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)")
```

```{r}
rownames(t1) <- NULL
kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2. 1,000 iterations were run with a burn in of 100. Missingness mechanism was MAR and P(miss) = 0. Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.") %>%
  kable_styling(font_size = 7) %>%
  add_header_above(c(" " = 2,"Class 1" = 3, "Class 2" = 3, "Class 3" = 3)) %>%
  group_rows(" ",1,8) %>%
  group_rows(" ",9,18) %>%
  group_rows(" ",19,22) %>%
  group_rows(" ",23,26) %>%
  group_rows(" ",27,27) %>%
  landscape()
```

>>>>

model_component <- c("MVSN","Regression","","","","","","","","","","",
                     "Multinom.","","","",
                     "Clustering")
params <- c("$\\beta_{11}$","$\\beta_{21}$","$\\beta_{31}$","$\\beta_{41}$",
            "$\\beta_{12}$","$\\beta_{22}$","$\\beta_{32}$","$\\beta_{42}$",
            "$\\alpha_{1}$","$\\alpha_{2}$", "$\\alpha_{3}$","$\\alpha_{4}$",
            "$\\delta_{11}$","$\\delta_{12}$",
            "$\\delta_{21}$","$\\delta_{22}$",
            "$\\pi_l$")

true_c1 <- c(beta_true_c1,alpha_true_c1,delta_true,pi_true[1])
true_c2 <- c(beta_true_c2,alpha_true_c2,delta_true,pi_true[2])
true_c3 <- c(beta_true_c3,alpha_true_c3,delta_true,pi_true[3])
est_c1 <- c(beta_est_c1,alpha_est_c1,delta_est,pi_est[1])
est_c2 <- c(beta_est_c2,alpha_est_c2,delta_est,pi_est[2])
est_c3 <- c(beta_est_c3,alpha_est_c3,delta_est,pi_est[3])

@@

t1 <- cbind(model_component,
            params,
            true_c1,est_c1,est_c2_noskew[-(9:18)])
colnames(t1) <- c("Component",
                  "Param.",
                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI) ")
```

```{r}
rownames(t1) <- NULL
kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2. 1,000 iterations were run with a burn in of 100. Missingness mechanism was MAR and P(miss) = 0. Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.") %>%
  kable_styling(font_size = 7) %>%
  add_header_above(c(" " = 2,"Class 1" = 3)) %>%
  group_rows(" ",1,8) %>%
  group_rows(" ",9,12) %>%
  group_rows(" ",13,16) %>%
  group_rows(" ",17,17)
```
